#**Peer-graded Assignment: Getting and Cleaning Data Course Project**
====================================================================
## By Abhinav

## *Loading and preprocessing the data*

Download and unzip the file

```{r Download_Unzip, echo=TRUE}
fileurl<-"https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
download.file(fileurl,"Dataset.zip")
unzip("Dataset.zip")
```

Read files in Data Frames

```{r LoadData1, echo=TRUE}
library(data.table)
library(dplyr)
library(plyr)
activity_labels<-read.table("activity_labels.txt")
features<-read.table("features.txt")
str(activity_labels)
str(features)
```

```{r LoadData2, echo=TRUE}
setwd("./train")
subject_train<-read.table("subject_train.txt")
X_train<-read.table("X_train.txt")
Y_train<-read.table("y_train.txt")
dim(subject_train)
dim(X_train)
dim(Y_train)
```
```{r LoadData3, echo=TRUE}
setwd("./Inertial_Signals")
body_acc_x_train<-read.table("body_acc_x_train.txt")
body_acc_y_train<-read.table("body_acc_y_train.txt")
body_acc_z_train<-read.table("body_acc_z_train.txt")

body_gyro_x_train<-read.table("body_gyro_x_train.txt")
body_gyro_y_train<-read.table("body_gyro_y_train.txt")
body_gyro_z_train<-read.table("body_gyro_z_train.txt")

total_acc_x_train<-read.table("total_acc_x_train.txt")
total_acc_y_train<-read.table("total_acc_y_train.txt")
total_acc_z_train<-read.table("total_acc_z_train.txt")
```

```{r LoadData4, echo=TRUE}
setwd("../")
setwd("../")
setwd("./test")

subject_test<-read.table("subject_test.txt")
X_test<-read.table("X_test.txt")
Y_test<-read.table("y_test.txt")
dim(subject_test)
dim(X_test)
dim(Y_test)
```

```{r LoadData5, echo=TRUE}
setwd("./Inertial_Signals")
body_acc_x_test<-read.table("body_acc_x_test.txt")
body_acc_y_test<-read.table("body_acc_y_test.txt")
body_acc_z_test<-read.table("body_acc_z_test.txt")

body_gyro_x_test<-read.table("body_gyro_x_test.txt")
body_gyro_y_test<-read.table("body_gyro_y_test.txt")
body_gyro_z_test<-read.table("body_gyro_z_test.txt")

total_acc_x_test<-read.table("total_acc_x_test.txt")
total_acc_y_test<-read.table("total_acc_y_test.txt")
total_acc_z_test<-read.table("total_acc_z_test.txt")
```

## *1. Merge the training and the test sets to create one data set.*

-Merges Train Data using cbind
-Merges Variables "Activity", "Subjects", 
-Feature Data,
-Inertial Signals (total_acc, body_acc, body_gyro in the order x,y,z)

```{r trainDataBind, echo=TRUE}
trainData<-cbind(Y_train,subject_train,X_train,body_acc_x_train,body_acc_y_train,body_acc_z_train,body_gyro_x_train,body_gyro_y_train,body_gyro_y_train,total_acc_x_train,total_acc_y_train,total_acc_z_train)
dim(trainData)
```
Similarly merge Test Data using cbind


```{r testDataBind, echo=TRUE}
testData<-cbind(Y_test,subject_test,X_test,body_acc_x_test,body_acc_y_test,body_acc_z_test,body_gyro_x_test,body_gyro_y_test,body_gyro_y_test,total_acc_x_test,total_acc_y_test,total_acc_z_test)
dim(testData)
```

Merge Train Data and Test Data using rbind and assign to totalData

```{r totalDataBind, echo=TRUE}
totalData<-rbind(trainData,testData)
dim(totalData)
```
## *2. Extracts only the measurements on the mean and standard deviation for each measurement*

-From the "features Data" Frame, extract those features as TRUE which has mean and standard deviation in it using grepl function
-Extract these columns from totalData and assign to totalData_mean_std 

```{r ExtractMeanStd, echo=TRUE}
totalData_mean_std<-totalData[,c(rep(FALSE,times=2),(grepl("std",features$V2)|grepl("mean",features$V2)),rep(FALSE,times=128*9))]
dim(totalData_mean_std)
```

## *3. Uses descriptive activity names to name the activities in the data set*

Merge "totalData" and "activity_labels" to get descriptive activity names


```{r ActivityLabels, echo=TRUE}
totalData_Activity<-merge(totalData,activity_labels,by.x=1,by.y=1,all=TRUE)
totalData_Activity[1:5,c(1,1716)]
```

## 4. Appropriately labels the data set with descriptive variable names. 

-Var1: ActivityCode
-Var2: SubjectCode
-Var3 to Var 563: Features
-Var564 to Var 1715: Inertial Signals
-Var1716: ActivityLabels

```{r LabelDataSet, echo=TRUE}
colnames(totalData_Activity)<-c("ActivityCode","SubjectCode",as.character(features$V2),paste("body_acceleration_x",1:128),paste("body_acceleration_y",1:128),paste("body_acc_z",1:128),paste("body_gyro_x",1:128),paste("body_gyro_y",1:128),paste("body_gyro_z",1:128),paste("total_acceleration_x",1:128),paste("total_acceleration_y",1:128),paste("total_acceleration_z",1:128),"ActivityLabels")
```

-Identify and rename duplicate variables
-Identify and rename variables containing special characters

```{r RefineName, echo=TRUE}
names(totalData_Activity)[duplicated(names(totalData_Activity))]<-paste(names(totalData_Activity)[duplicated(names(totalData_Activity))],rep(1:2,each=14,times=3))
names(totalData_Activity)<-make.names(names(totalData_Activity),unique=TRUE)
colnames(totalData_Activity)
```

## *5. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.*
-Assign "totalData_activity" with removed last column to "totalData1" 
-melt Data with id as ActivityCode and SubjectCode and rest of the measurments as Variables
-dcast the melted Data with mean of melted columns on ActivityCode+SubjectCode and save in another tidy data "finalData"

```{r CalcAverage, echo=TRUE}
totalData1<-totalData_Activity[,1:1715]
meltData<-melt(totalData1,id=c("ActivityCode","SubjectCode"),measure.vars=3:1715)
finalData<-dcast(meltData,ActivityCode+SubjectCode~variable,mean)
```
Write the data to file "finalTidyData.txt" with row.names = FALSE

```{r WriteFile, echo=TRUE}
write.table(finalData,file="finalTidyData.txt",row.names=FALSE)
```


## *End of File*